Processing Code for Populations of Zips

### 2023-04-18 code/processing_code/processingcode2.qmd
Population sums by zip

---
title: "Processing script 2"
author: "Jacob Matta"
date: "2023-01-03"
output: html_document
---
# Processing script 2: Understanding the rawdata and data source

## Determining the count of residents in GA with at least one dose of the covid-19 vaccines

The count of residents in GA with at least one dose of the Covid-19 vaccines as of Feb 26th, 2023 is 6,935,491. This was found from the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th/

## This processing script will look to see how the raw data used for this analysis compares to the overall number of 6,935,491 as of Feb 26th, 2023.

The overall sum of residents in GA with at least one dose of the covid-19 vaccines as of Feb 26th, 2023 should equal 6,935,491

### Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(here) #used for setting a path 
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

### Raw Data loading

```{r}
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)
```

### Codebook loading

```{r}
#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Cleaning

```{r}
#dropping columns that i don't need
#keeping county name to group data together 
Feb_26 <- Feb_26[-c(1:3)]

#removing all of the population data 
Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
Feb_26[-c(1)]

#removing the negative values 
Feb_26 [Feb_26 < 0] <- 0

@@

### Summing the vaccinations in each column

```{r}
Feb_26_sum_by_col = colSums(Feb_26[ , c(2:91)])     
Feb_26_sum_by_col
```

### Turning the sums by column into a tibble that puts all of the sums of each column into one column

```{r}
Feb_26_sum_tib = tibble(Feb_26_sum_by_col)
Feb_26_sum_tib
```

### Summing all of the vaccinations together

```{r}
Overall_sum = sum(Feb_26_sum_tib) 
Overall_sum
```

### Thoughts

The overall total of residents in GA with at least one dose of the covid-19 vaccines is 5,654,635 which is more than 1 million should of the total about of GA residents with at least one dose of a Covid-19 vaccine that the Georgia Department of Public Health Vaccine Distribution Dashboard reported on Feb, 26th, 2023.

The data that I am using for this project possibly excludes certain vaccines brands (eg. Pfizer, Moderna, Johnson & Johnson) whereas the overall vaccination count represented on the website includes all vaccination brands. This is a question I will attempt to figure out. For now, it is my understanding that there are 5,654,635 Ga residents by census tract who currently have one dose of the Covid-19 vaccine. The brands of Covid-19 vaccines included in the census tract data has yet to be specified.

### Explanation for difference Covid-19 Vaccination totals

1.  The data I downloaded is by census tract

2.  The data displayed on the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th is by county level which does not align with census tracts

3.  Going forward for details regarding the raw data you should go to arcGIS insights dashboard for Covid-19 vaccinations among GA Residents with at least one dose found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995)

4.  The data dashboard for Covid-19 vaccinations among GA Residents with at least one dose (link above) states that the data downloaded by census tract will not add up to the county level and state level data

### Moving forward
If the vaccinations are not rates (eg. per 100,000) why are there negative values and why does it state that the where "-1 is indicated, either the numerator is less than five or the denominator is less than 25".
-   What is the numerator and what is the denominator?

>>>>

---
title: "Processing Code for Populations of Zips"
output: 
  html_document
---
# Processing Section for Populations of Target Zip Codes

## Processing 1: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of September

### Load needed packages. make sure they are installed.
```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_POP_Sept <- filter(Sept_17, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_POP_Sept [Dougherty_POP_Sept < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Dougherty_POP_Sept <- Dougherty_POP_Sept  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

@@

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Dougherty_POP_Sept<- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

```{r}
#POP total Asian Residents
Dougherty_POP_Sept$total_asian_POP <- Dougherty_POP_Sept$total_masian_POP + Dougherty_POP_Sept$total_fasian_POP

#POP total for African American Residents 
Dougherty_POP_Sept$total_black_POP <- Dougherty_POP_Sept$total_mblack_POP + Dougherty_POP_Sept$total_fblack_POP

#POP total for Other Race Residents
Dougherty_POP_Sept$total_orace_POP <- Dougherty_POP_Sept$total_morace_POP + Dougherty_POP_Sept$total_forace_POP

#POP total for White Residents
Dougherty_POP_Sept$total_white_POP <- Dougherty_POP_Sept$total_mwhite_POP + Dougherty_POP_Sept$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_POP_Sept_sums <- Dougherty_POP_Sept[,c(1:4,185:204)]  
```

### Filtering the data by zip code

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Sept_POP <-filter(Dougherty_POP_Sept_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Sept_POP <- filter(Dougherty_POP_Sept_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Sep_POP <- filter(Dougherty_POP_Sept_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

### Change the column names for 31701 Sept

```{r}
colnames(Zip_31701_Sept_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")

#change column names for 31705 Sept  
colnames(Zip_31705_Sept_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")

#change column names for 31707 Sept  
colnames(Zip_31707_Sep_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")

### Deleting excess environment objects and saving data as processed_data_sept
```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_POP_Sept,Sept_17, codebook, Dougherty_POP_Sept_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Sept.rds")

saveRDS(Zip_31701_Sept_POP, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Sept.rds")

saveRDS(Zip_31705_Sept_POP, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Sept.rds")

saveRDS(Zip_31707_Sep_POP, file = save_data_location3)

## Processing 2: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of October 
## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Octarating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_POP_Oct <- filter(Oct_28, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_POP_Oct [Dougherty_POP_Oct < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Dougherty_POP_Oct <- Dougherty_POP_Oct  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Dougherty_POP_Oct<- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

```{r}
#POP total Asian Residents
Dougherty_POP_Oct$total_asian_POP <- Dougherty_POP_Oct$total_masian_POP + Dougherty_POP_Oct$total_fasian_POP

#POP total for African American Residents 
Dougherty_POP_Oct$total_black_POP <- Dougherty_POP_Oct$total_mblack_POP + Dougherty_POP_Oct$total_fblack_POP

#POP total for Other Race Residents
Dougherty_POP_Oct$total_orace_POP <- Dougherty_POP_Oct$total_morace_POP + Dougherty_POP_Oct$total_forace_POP

#POP total for White Residents
Dougherty_POP_Oct$total_white_POP <- Dougherty_POP_Oct$total_mwhite_POP + Dougherty_POP_Oct$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_POP_Oct_sums <- Dougherty_POP_Oct[,c(1:4,185:204)]  
```

### Filtering the data by zip code

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Oct_POP <-filter(Dougherty_POP_Oct_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Oct_POP <- filter(Dougherty_POP_Oct_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Oct_POP <- filter(Dougherty_POP_Oct_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

### Change the column names for 31701 Oct

```{r}
colnames(Zip_31701_Oct_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")

#change column names for 31705 Oct  
colnames(Zip_31705_Oct_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")

#change column names for 31707 Oct  
colnames(Zip_31707_Oct_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")

### Deleting excess environment objects and saving data as processed_data_Oct
```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_POP_Oct,Oct_28, codebook, Dougherty_POP_Oct_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Oct.rds")

saveRDS(Zip_31701_Oct_POP, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Oct.rds")

saveRDS(Zip_31705_Oct_POP, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Oct.rds")

saveRDS(Zip_31707_Oct_POP, file = save_data_location3)

## Processing 3: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of November

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Novarating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_POP_Nov <- filter(Nov_24, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_POP_Nov [Dougherty_POP_Nov < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Dougherty_POP_Nov <- Dougherty_POP_Nov  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Dougherty_POP_Nov<- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

```{r}
#POP total Asian Residents
Dougherty_POP_Nov$total_asian_POP <- Dougherty_POP_Nov$total_masian_POP + Dougherty_POP_Nov$total_fasian_POP

#POP total for African American Residents 
Dougherty_POP_Nov$total_black_POP <- Dougherty_POP_Nov$total_mblack_POP + Dougherty_POP_Nov$total_fblack_POP

#POP total for Other Race Residents
Dougherty_POP_Nov$total_orace_POP <- Dougherty_POP_Nov$total_morace_POP + Dougherty_POP_Nov$total_forace_POP

#POP total for White Residents
Dougherty_POP_Nov$total_white_POP <- Dougherty_POP_Nov$total_mwhite_POP + Dougherty_POP_Nov$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
